\speculake: Hiding Malware in Speculative Execution

### 2018-08-06 paper/discussion.tex
add discussion on defenses

\subsection{Defenses}
We now address possible defenses to detecting and reverse engineering malware that uses
\speculake. 
\subsubsection{Implemented Mitigations}
Multiple patches and micro-code updates have been developed to mitigate Spectre
vulnerabilities, however, none of these entirely prevent \speculake malware from
working, as they are generally not designed to protect programs that willingly use
Spectre against themselves.

@@

%\paragraph{Red-Pills}
%Hardware minutia have also been used to identify and fingerprint execution
%environments---in malware this is known as a red-pill~\cite{red-pill} and is
%typically used by malware to detect when it is under emulation or a debugging
%environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
%paleari2009fistful} attempting to circumvent malware detectors. To address
%red-pills, systems often employ measures to detect them.

@@

%\paragraph{Operating System Defense}
%Hardware could also provide the operating system with a privileged instruction
%that gives it more control over the interaction between programs. For example,
%the OS could use a privileged instruction to copy out the Branch History Buffer
%(BHB) state when it context switches between processes, restoring it when the
%process is restored. This would prevent cross-process interactions with the
%branch predictor, and prevent multi-process \speculake malware. However, it does
%not prevent a single-process \speculake, and it does not fully address the
%Spectre vulnerability.
%

%\subsubsection{Cache Poisoning}
%As mentioned in Section~\ref{subsubsec:malware} a vital component of \speculake
%is the ability to perform cache timings. Any mechanism that modifies the cache
%can affect the success of \speculake-like attacks.
%
%Drastic methods such as completely poisoning the cache will hinder this attack
%while also forcing almost all loads to result in a cache miss, effectively
%nullifying the point of cache.
%
%However, more conservative methods such as randomly loading values into cache
%can have a negative affect on \speculake-like attacks (causing the cache timing
%to predict incorrect values) while not having too detrimental of an affect on
%programs as a whole.
%
%In response to this method of defense \speculake-like attacks can modify the
%number of iterations and confidence thresholds necessary to determine a value.
%As mentioned in Section~\ref{subsec:impl-turing} our experiments found that it
%was only necessary to perform 50 iterations of our \texttt{target\_fn} in order
%to gain 80\% confidence in the value returned. While only 50 iterations are
%necessary, increasing this value to higher levels can lead to a significantly
%higher confidence with very little performance detriment (especially in the
%case of AES decryption).

>>>>

\subsection{Defenses}
We now address possible defenses to detecting and reverse engineering malware
that uses \speculake.

\subsubsection{Implemented Mitigations}
Multiple patches and micro-code updates have been developed to mitigate Spectre
vulnerabilities, however, none of these entirely prevent \speculake malware from
working, as they are generally not designed to protect programs that willingly
use Spectre against themselves.

@@

%\paragraph{Red-Pills}
%Hardware minutia have also been used to identify and fingerprint execution
%environments---in malware this is known as a red-pill~\cite{red-pill} and is
%typically used by malware to detect when it is under emulation or a debugging
%environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
%paleari2009fistful} attempting to circumvent malware detectors. To address
%red-pills, systems often employ measures to detect them.
\paragraph{Anti-Virus Detectors}
We verified that numerous modern Anti-Virus technologies were unable to detect
and flag \speculake malware. We examined ClamAV, BitDefender and rkhunter and
found that they focus on signature and string detection, examining if potential
malware make library calls, though BitDefender does use their own proprietary
adaptable unpacker to support detection of ``generic
unpacking''~\cite{BitDefender}. Modern anti-virus detectors but do not flag
programs that directly make syscalls or have the capability of making arbitrary
syscalls. We note that this is the state of modern commercial anti-virus
technology but does not mean that \speculake is undetectable.

@@

\subsubsection{Reverse-engineering triggers}
Once \speculake malware has been detected analysts attempting to
reverse-engineer the effect of the malware can take a similar path as to its
authors. Specifically the task of examining the real-world execution path of the
malware and comparing this with the indirect and direct jumps conducted by
potential trigger programs to find a match. We note that while there may be a
large number of potential trigger programs the real world execution pathway must
match the malware's very closely in order to influence the malware successfully.
However this method will not be useful against the speculative store bypass
version of \speculake, as a secondary trigger program is unnecessary there.
